On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-06 main.tex
Updates from Overleaf

\begin{theorem}
  There exists a hypersurface $\D \subset \mathbb{C}^{n^2}$ of degree
  at most $\td$ such that, if $\mA$ is chosen from $\C^{n^2}-\Delta$
  then $\mA$ is invertible and $f^\mA$ satisfies $\textbf{H}_i$ for
  all $i \in \{1,\hdots,n-1\}.$
\end{theorem}
%% %
%% \begin{corollary} 
%%   Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1}
%%   D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen
%%   randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i
%%   \in \{1,\hdots,n\}$ is at least $1-\epsilon.$
%% \end{corollary}

@@

is integral.
%
%
%
%
\section{Proof of Theorem 2.1}
%
Recall Theorem 3.1: \textit{Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$} 
%
\begin{proof}
Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
\[
A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
\]
then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
\[
\Delta = \bigcup_{i=1}^3 \D_i,
\]
and note that 
\[
\deg \D \leq  \sum_{i=1}^3 \deg \D_i.
\]
\end{proof}
%
Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
%
\begin{proof}
\[
Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
\]
\end{proof}
%
%
%
%
\section{Bit complexity}
%
Recall Theorem 1.1: Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

  There exists a randomized algorithm that takes $\Gamma, d,$ and $s$
  as input and produces a zero-dimensional parameterization of a set
  that contains at least one point in each connected component of
  $V(f) \cap \R^n$ with probability at least $1/2$. Otherwise the
  algorithm either produces a subset of the points or FAIL. In any
  case, the algorithm uses
  \[
  O^{\sim}(k(Lb+d^{2n}(s+d)(L+d)))
  \]
  bit operations, with $k \geq -\lg(1-2^{(-1/n)})$. The polynomials in
  the output have degree at most $d^n,$ and height $
  O^{\sim}((s+d)d^n).$

Recall as well that the algorithm can error on one side by producing a proper subset of the points. Running the algorithm $\kt$ times gives a list of $n$ zero-dimensional parameterizations, among which the highest cardinality output includes at least one point in each connected component with probability at least $1-(1/2)^{\kt}.$
%
\subsection{The Algorithm}
%
For a matrix $A \in \GL(n,\Q),$ let 
\[
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
\]
where $b_i$ is the i-th row of $A^{-1}$.
%
\begin{algorithm}[hbt!]
\KwIn{a strait line program $\Gamma$ that computes $f \in \Q[X_1,\hdots,X_n],$ and $d=\deg(f)$}
\KwOut{$n$ zero-dimensional parameterizations, whose reunion includes at least one point in each connected component of $V(f) \cap \R^n$}
\nl Choose $S \subset \mathbb{Q}$ and $A \in S^{n^2}$ with $|S|\geq \epsilon^{-1}d^{\td}$, where $\epsilon < 5/21$\;
\nl Compute $f$ and apply the change of variables to obtain $f(A x)=f^A(x)$\;
\nl Use \cite{partialDerComputation} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;
    \caption{{Main Algorithm} \label{}}
\nl Randomly choose a point $p_{n-1}=(\alpha_1,\hdots,\alpha_{n-1}) \in \Q^{n-1}~(p_0$?)\; 
\nl \For{$i\gets1$ \KwTo $n$}{
\nl Run \cite[Algorithm 2]{SH} $k > -\lg(1-2^{(-1/n)})$ times with input 
$\{Y_1-\alpha_1,\hdots,Y_{i-1}-\alpha_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \} \textrm{ and } d$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 6 \;
} 
\nl \Return $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$.
\end{algorithm}

%
\subsection{Proof of Theorem 1.1}
%
\textbf{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
%
\begin{enumerate}
    \item By our assumption on the size of $S$ we know that $|S| = \log (\epsilon^{-1}D^{\td}) \leq \td.$ Thus for $a_{ij} \in S$ we know that $|a_{ij}| \leq \td.$
    \item 
    \begin{enumerate}[label=\roman*.]
        \item $O(L)$ operations are required to compute $f$ from its straight line program $\Gamma.$
        \item After applying the change of variables obtaining $f(A x)=f^A(x)$ the height grows at most by a factor of
        \begin{align*}
        \log(n!\cdot \max_{\frac{1\leq i\leq n}{1\leq j\leq n}}\{a_{ij}\}) &= 
        O(n\log n + n\log(d+1)) \\
        &= 
        O^\sim(\log(d+1)).
        \end{align*}
    \end{enumerate}
    \item After computing the partial derivatives the height grows by at most another factor of $\log d$. Furthermore, preparing the straight line program for the gradient of $f$ does not introduce large integers. The integers in $\grad (f)$ are of the same magnitude as in $\Gamma$.
    \item $\td$
    \item 
    After applying \cite[Algorithm 2]{SH} $kn$ times, we obtain the
    following upper bound on the total number of bit operations required by the algorithm:
    \[
    O^\sim(kn(Lb+d^{2n}(s+d)(L+d))),
    \]
    where it also follows from \cite[Algorithm 2]{SH} that the polynomials in the output have degree at most $d^n,$ and height at most
    \[
    O^{\sim}((s+d)d^n).
    \]
\end{enumerate}
%
\par \noindent 
\textbf{Probability of success:} Note that when $\epsilon < 5/21$ (as stated in step 1 of Algorithm 1), then
\[
\pr[A \not \in \D] = 1-\epsilon > \frac{16}{21},
\]
so that after a single run of \cite[Algorithm 2]{SH}, the probability of success (ensuring all genericity properties and obtaining all points) is at least 
\[
\left( \frac{16}{21} \right) \left( \frac{21}{32} \right) = \frac{1}{2}.
\]
Thus, running \cite[Algorithm 2]{SH} $k$ times gives a list of outputs among which the highest cardinality set includes all points with probability at least $1-(1/2)^k$. We repeat this $n$ times to obtain $n$ zero-dimensional parameterizations. Thus, the probability that we have every point in each parameterization is $(1-(1/2)^k)^n$. We want 
\[
(1-(1/2)^k)^n>\frac{1}{2},
\]
which holds if and only if 
\[
k > -\lg(1-2^{(-1/n)}),
\]
thus proving Theorem 1.1. We can boost our success probability by running the main algorithm $\kt$ times, so that we obtain parameterizations whose reunion includes at least one point in each connected component of $V(f) \cap \R^n$ with probability at least $1 - (1/2)^{\kt}$. 
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
\endinput
%%
%% End of file `sample-xelatex.tex'.

>>>>

\begin{theorem}
  There exists a hypersurface $\D \subset \mathbb{C}^{n^2}$ of degree
  at most $\td$ such that, if $\mA$ is chosen from $\C^{n^2}-\Delta$
  then $\mA$ is invertible and $f^\mA$ satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$ for
  all $i \in \{1,\hdots,n-1\}.$
\end{theorem}
%% %
%% \begin{corollary} 
%%   Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1}
%%   D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen
%%   randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i
%%   \in \{1,\hdots,n\}$ is at least $1-\epsilon.$
%% \end{corollary}

@@

is integral.
%
%
%
%
\section{Bit complexity}
%
Recall Theorem 1.1: Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, satisfies
  $\deg(f) \leq d, \htt(f) \leq s$, and that $V(f) \subset \C^n$ is
  smooth.  Assume also that $f$ is given by a straight-line program
  $\Gamma$ of size $L$ with integer constants of height at most $b.$

  There exists a randomized algorithm that takes $\Gamma, d,$ and $s$
  as input and produces a zero-dimensional parameterization of a set
  that contains at least one point in each connected component of
  $V(f) \cap \R^n$ with probability at least $1/2$. Otherwise the
  algorithm either produces a subset of the points or FAIL. In any
  case, the algorithm uses
  \[
  O^{\sim}(nk(Lb+d^{2n}(s+d)(L+d)))
  \]
  bit operations, with $k \geq -\lg(1-2^{(-1/n)})$. The polynomials in
  the output have degree at most $d^n,$ and height $
  O^{\sim}((s+d)d^n).$

Recall as well that the algorithm can error on one side by producing a proper subset of the points. Running the algorithm $\kt$ times gives a list of $n$ zero-dimensional parameterizations, among which the highest cardinality output includes at least one point in each connected component with probability at least $1-(1/2)^{\kt}.$
%
\subsection{The Algorithm}
%
For a matrix $A \in \GL(n,\Q),$ let 
\[
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
\]
where $b_i$ is the i-th row of $A^{-1}$.
%
\begin{algorithm}[hbt!]
\KwIn{a strait line program $\Gamma$ that computes $f \in \Q[X_1,\hdots,X_n],$ and $d=\deg(f)$}
\KwOut{$n$ zero-dimensional parameterizations, whose reunion includes at least one point in each connected component of $V(f) \cap \R^n$}
\nl Choose $S \subset \mathbb{Q}$ and $A \in S^{n^2}$ with $|S|\geq \epsilon^{-1}d^{\td}$, where $\epsilon < 5/21$\;
\nl Compute $f$ and apply the change of variables to obtain $f(A x)=f^A(x)$\;
\nl Use \cite{partialDerComputation} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;
    \caption{{Main Algorithm} \label{}}
\nl Choose $T \subset \mathbb{Q}$ and $\alpha_{n-1}=(\alpha_1,\hdots,\alpha_{n-1}) \in T^{n-1}$ with $|T| \geq \epsilon^{-1}d^{\td}$\; 
\nl \For{$i\gets1$ \KwTo $n$}{
\nl Run \cite[Algorithm 2]{SH} $k > -\lg(1-2^{(-1/n)})$ times with input 
$\{Y_1-\alpha_1,\hdots,Y_{i-1}-\alpha_{i-1},f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \} \textrm{ and } d$\;
\nl Let $\mathscr{Q}_i$ be the highest cardinality zero-dimensional parameterization returned in step 6 \;
} 
\nl \Return $[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$.
\end{algorithm}
\clearpage 
%
\subsection{Proof of Theorem 1.1}
%
Recall Theorem 2.1: \textit{  There exists a hypersurface $\D \subset \mathbb{C}^{n^2}$ of degree at most $\td$ such that, if $\mA$ is chosen from $\C^{n^2}-\Delta$ then $\mA$ is invertible and $f^\mA$ satisfies $\textbf{H}_i$ and $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n-1\}.$} 
%
\begin{proof}
Assume that $f\in \Q[X_1,\hdots,X_n]$ has degree at most $D$ and satisfies $\textbf{H}.$ By Theorem 5.1, there exists a Zariski closed set $\D_1\subset \C^{n^2}$ with the property that, if $A \in \C^{n^2}-\D_1$ then, for $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1), \textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Likewise, by Theorem 6.1, we obtain $\D_2 \subset \C^{n^2}$ with the property that if $A \in \C^{n^2}-\D_2$ then $f^A$ satisfies $\textbf{H}_i^{'}(4), \textbf{H}_i^{'}(5),$ and $\textbf{H}_i^{'}(6).$ By Theorem 7.1, $\D_3 \subset \C^{n^2}$ exists with $f^A$ satisfying $\textbf{H}_i^{'}(7)$ when $A \in \C^{n^2}-\D_3$. Clearly then, if 
\[
A \in \C^{n^2} - \bigcup_{i=1}^3 \D_i
\]
then $f^A$ satisfies  $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$ Take 
\[
\Delta = \bigcup_{i=1}^3 \D_i,
\]
and note that 
\[
\deg \D \leq  \sum_{i=1}^3 \deg \D_i.
\]
\end{proof}
%
Recall Corollary 3.2: \textit{Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$}
%
\begin{proof}
\[
Pr[A \in \Delta] \leq \frac{\deg \Delta}{|S|} \leq \frac{D^{\td}}{\epsilon^{-1}D^{\td}} = \epsilon.
\]
\end{proof}
\textbf{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
%
\begin{enumerate}
    \item We assume that $S=\{1,2,\hdots,\epsilon^{-1}\td\}$. Therefore, the height of $a_{ij} \in S$ is bounded by $\log \td -\log \epsilon.$
    \item 
    \begin{enumerate}[label=\roman*.]
        \item $O(L)$ operations are required to compute $f$ from its straight line program $\Gamma.$
        \item After applying the change of variables obtaining $f(A x)=f^A(x)$ the height grows at most by a factor of
        \begin{align*}
        \log(n!\cdot \max_{\frac{1\leq i\leq n}{1\leq j\leq n}}\{a_{ij}\}) &= 
        O(n\log n + n\log(d+1)) \\
        &= 
        O^\sim(n\log(d)- \log \epsilon).
        \end{align*}
    \end{enumerate}
    \item After computing the partial derivatives the height grows by at most another factor of $\log d$. Furthermore, preparing the straight line program for the gradient of $f$ does not introduce large integers. The integers in $\grad (f)$ are of the same magnitude as in $\Gamma$.
    \item $\td$
    \item 
    After applying \cite[Algorithm 2]{SH} $kn$ times, we obtain the
    following upper bound on the total number of bit operations required by the algorithm:
    \[
    O^\sim(kn(Lb+d^{2n}(s+d)(L+d))),
    \]
    where it also follows from \cite[Algorithm 2]{SH} that the polynomials in the output have degree at most $d^n,$ and height at most
    \[
    O^{\sim}((s+d)d^n).
    \]
\end{enumerate}
%
\par \noindent 
\textbf{Probability of success:} Note that when $\epsilon < 5/21$ (as stated in step 1 of Algorithm 1), then
\[
\pr[A \not \in \D] = 1-\epsilon > \frac{16}{21},
\]
so that after a single run of \cite[Algorithm 2]{SH}, the probability of success (ensuring all genericity properties and obtaining all points) is at least 
\[
\left( \frac{16}{21} \right) \left( \frac{21}{32} \right) = \frac{1}{2}.
\]
Thus, running \cite[Algorithm 2]{SH} $k$ times gives a list of outputs among which the highest cardinality set includes all points with probability at least $1-(1/2)^k$. We repeat this $n$ times to obtain $n$ zero-dimensional parameterizations. Thus, the probability that we have every point in each parameterization is $(1-(1/2)^k)^n$. We want 
\[
(1-(1/2)^k)^n>\frac{1}{2},
\]
which holds if and only if 
\[
k > -\lg(1-2^{(-1/n)}),
\]
thus proving Theorem 1.1. We can boost our success probability by running the main algorithm $\kt$ times, so that we obtain parameterizations whose reunion includes at least one point in each connected component of $V(f) \cap \R^n$ with probability at least $1 - (1/2)^{\kt}$. 
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
\endinput
%%
%% End of file `sample-xelatex.tex'.
